---
title: "Visualisations for EMS Response Time as a function of different predictors"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(plotly)
library(maps)
library(mapdata)
library(viridis)
library(ggthemes)
library(lubridate)
library(patchwork)
library(kableExtra)
library(magick)
library(jpeg)
library(grid)

```

Here we showcase a dhasboard of the various plots that led us to come up with the analysis.

```{r data_import_tidying, echo = FALSE}
load('data/finaldat.RData')
load("data/incident_zip_coor.RData")

finaldat = 
  finaldat %>%
  mutate(season = 
           ifelse(incident_month %in% 9:11, "Fall",
           ifelse(incident_month %in% c(12,1,2), "Winter",
           ifelse(incident_month %in% 3:5, "Spring", "Summer"))), 
         hour_of_day = 
           ifelse(hour %in% 6:11, "morning",
           ifelse(hour %in% 12:17, "afternoon",
           ifelse(hour %in% 18:23, "night","dawn"))), 
         over_8min = ifelse(response_time > 8, "8min+", "8min-"),
         prcp_ctg = 
           if_else(prcp == 0, "no_prcp",
           if_else((prcp > 0 & prcp <= 25), "low", "high")),
         snow_ctg = 
           if_else(snow == 0, "no_snow",
           if_else((snow > 0 & snow <= 50), "low", "high")),
         over_8min = fct_relevel(over_8min, "8min-")) %>% 
  mutate(prcp_ctg = fct_relevel(prcp_ctg, "no_prcp"),
         snow_ctg = fct_relevel(snow_ctg, "no_snow"),
         season = fct_relevel(season, "Spring"),
         hour_of_day = fct_relevel(hour_of_day, "night"))

fit_logistic =
  glm(over_8min ~ season + hour_of_day + snow_ctg + prcp_ctg, 
      family = binomial(), data = finaldat)

OR_total_df = 
fit_logistic %>% 
  broom::tidy() %>% 
  mutate(OR = exp(estimate),
         lower_CI_OR = exp(estimate - 1.96*std.error),
         upper_CI_OR = exp(estimate + 1.96*std.error)) %>% 
  select(term, OR, lower_CI_OR, upper_CI_OR) %>%
  as.tibble() %>% 
  mutate(ctg = c("intercept", 
                 "Season", "Season", "Season", 
                 "Hour of the day", "Hour of the day", "Hour of the day",
                 "Snow", "Snow",
                 "Rain", "Rain"),
         sub_ctg_chr = c("intercept",
                     "Fall", "Summer", "Winter",
                     "Afternoon", "Dawn", "Morning",
                     "50(mm)+", "0-50(mm)",
                     "25(mm)+", "0-25(mm)"),
         sub_ctg = c("1", "2", "1", "3", "8", "6", "7", "5", "4", "5", "4")) %>%   
  filter(ctg != "intercept")  

adj = .2 # This is used in position_nudge to move the dots

#ny = map_data("county", "new york")

#ny_1 = 
#  ny %>% 
#  filter(subregion %in% c("kings", "bronx", "richmond", "new york", "queens"))

#incident_zip_coor = 
#  incident_zip_coor %>%
#  group_by(zip_code) %>% 
#  mutate(response_time = as.numeric(response_time)) %>%
#  mutate(mean_res_time = mean(response_time))
```

Column {data-width=500}
-----------------------------------------------------------------------

### Variation of Mean Response Time over different hours of days through the year, per season

```{r}
hour_eda =
  finaldat %>%
  group_by(hour, season) %>% 
  ggplot(aes(x = hour, y = response_time, color = season)) + 
  geom_smooth(se = FALSE) +
  scale_x_continuous(breaks = c(6, 12, 18),
                     labels = c("6am", "12pm", "18pm")) +
  geom_vline(xintercept = c(6, 12, 18), color = "darkred") +
  labs(
    x = " ", 
    y = "Mean response time (min)"
    ) +
   viridis::scale_color_viridis(
    name = "Season",
    discrete = TRUE) +
  theme(plot.title = element_text(size = 13),
        axis.title.y = element_text(size = 10),
        axis.text.x = element_text(size = 12),
        legend.position = "None") 
  
ggplotly(hour_eda)
```

Column {data-width=500}
-----------------------------------------------------------------------

### Odds Ratio with respect to weather, rain and hour of day.

```{r}
OR =
  ggplot(OR_total_df, aes(x = OR, y = ctg, color = sub_ctg, 
                        label = sub_ctg_chr)) +
  geom_vline(aes(xintercept = 1), size = 1, linetype = "dashed") +
  geom_vline(aes(xintercept = 1.5), size = .6, linetype = "dashed") +
  geom_errorbarh(data = filter(OR_total_df, sub_ctg == "1"), 
                 aes(xmax = upper_CI_OR, xmin = lower_CI_OR), 
                 size = .5, height = .1, 
                 color = "gray50", 
                 position = position_nudge(y = adj)) +
  geom_text(data = filter(OR_total_df, sub_ctg == "1"),
            aes(label = sub_ctg_chr), size = 6, colour = "seagreen4", 
            nudge_x = -.1, nudge_y = +.2, check_overlap = TRUE) +
  geom_point(data = filter(OR_total_df, sub_ctg == "1"), 
             size = 6, color = "seagreen4",
             position = position_nudge(y = adj)) +
  geom_errorbarh(data = filter(OR_total_df, sub_ctg == "2"), 
                 aes(xmax = upper_CI_OR, xmin = lower_CI_OR), 
                 size = .5, height = .1, 
                 color = "gray50") +
  geom_text(data = filter(OR_total_df, sub_ctg == "2"),
            aes(label = sub_ctg_chr), size = 6, colour = "indianred4", 
            nudge_x = -.1, nudge_y = 0, check_overlap = TRUE) +
  geom_point(data = filter(OR_total_df, sub_ctg == "2"), 
             size = 6, color = "indianred4") +
  geom_errorbarh(data = filter(OR_total_df, sub_ctg == "3"), 
                 aes(xmax = upper_CI_OR, xmin = lower_CI_OR), 
                 size = .5, height = .1, 
                 color = "gray50", position = position_nudge(y = -adj)) +
  geom_text(data = filter(OR_total_df, sub_ctg == "3"),
            aes(label = sub_ctg_chr), size = 6, colour = "lavenderblush4", 
            nudge_x = -.1, nudge_y = -.2, check_overlap = TRUE) +  
  geom_point(data = filter(OR_total_df, sub_ctg == "3"), 
             size = 6, color = "lavenderblush4",
             position = position_nudge(y = -adj)) +
  geom_errorbarh(data = filter(OR_total_df, sub_ctg == "4"), 
                 aes(xmax = upper_CI_OR, xmin = lower_CI_OR), 
                 size = .5, height = .1, 
                 color = "gray50", position = position_nudge(y = +.1)) +
  geom_text(data = filter(OR_total_df, sub_ctg == "4"),
            aes(label = sub_ctg_chr), size = 6, colour = "skyblue2", 
            nudge_x = -.15, nudge_y = +.1, check_overlap = TRUE) +
  geom_point(data = filter(OR_total_df, sub_ctg == "4"), 
             size = 6, color = "skyblue2",
             position = position_nudge(y = +.1)) +
  geom_errorbarh(data = filter(OR_total_df, sub_ctg == "5"), 
                 aes(xmax = upper_CI_OR, xmin = lower_CI_OR), 
                 size = .5, height = .1, 
                 color = "gray50", position = position_nudge(y = -.1)) +
  geom_text(data = filter(OR_total_df, sub_ctg == "5"),
            aes(label = sub_ctg_chr), size = 6, colour = "skyblue4", 
            nudge_x = -.16, nudge_y = -.1) +
  geom_point(data = filter(OR_total_df, sub_ctg == "5"), 
             size = 6, color = "skyblue4", 
             position = position_nudge(y = -.1)) +
 geom_errorbarh(data = filter(OR_total_df, sub_ctg == "6"), 
                 aes(xmax = upper_CI_OR, xmin = lower_CI_OR), 
                 size = .5, height = .1, 
                 color = "gray50", 
                 position = position_nudge(y = adj)) +
  geom_text(data = filter(OR_total_df, sub_ctg == "6"),
            aes(label = sub_ctg_chr), size = 6, colour = "seagreen4", 
            nudge_x = -.14, nudge_y = +.2, check_overlap = TRUE) +
  geom_point(data = filter(OR_total_df, sub_ctg == "6"), 
             size = 6, color = "seagreen4",
             position = position_nudge(y = adj)) +
  geom_errorbarh(data = filter(OR_total_df, sub_ctg == "7"), 
                 aes(xmax = upper_CI_OR, xmin = lower_CI_OR), 
                 size = .5, height = .1, 
                 color = "gray50") +
  geom_text(data = filter(OR_total_df, sub_ctg == "7"),
            aes(label = sub_ctg_chr), size = 6, colour = "indianred4", 
            nudge_x = -.13, nudge_y = 0, check_overlap = TRUE) +
  geom_point(data = filter(OR_total_df, sub_ctg == "7"), 
             size = 6, color = "indianred4") +
  geom_errorbarh(data = filter(OR_total_df, sub_ctg == "8"), 
                 aes(xmax = upper_CI_OR, xmin = lower_CI_OR), 
                 size = .5, height = .1, 
                 color = "gray50", position = position_nudge(y = -adj)) +
  geom_text(data = filter(OR_total_df, sub_ctg == "8"),
            aes(label = sub_ctg_chr), size = 6, colour = "lavenderblush4", 
            nudge_x = -.13, nudge_y = -.2, check_overlap = TRUE) +  
  geom_point(data = filter(OR_total_df, sub_ctg == "8"), 
             size = 6, color = "lavenderblush4",
             position = position_nudge(y = -adj)) +
  scale_x_continuous(breaks = c(1, 1.5),
                     limits = c(.7, 1.8)) +
  labs(x = "Odds Ratio",
       y = " ") +
  theme_bw() 

#OR + theme(panel.grid.minor = element_blank(),
#        axis.title.x = element_text(size = 16),
#        axis.text.x = element_text(size = 16),
#        axis.text.y = element_text(size = 20),
#        legend.position = "None") 



ggplotly(OR)
```
<!--

Column {data-width=500}
-----------------------------------------------------------------------

### Mean Response Time Distribution over all of NYC -- Visualised.


#map_plot = 
#  ggplot() + 
#  geom_polygon(data = ny_1, aes(x = long, y = lat, group = group, fill = subregion)) +
#  geom_point(data = incident_zip_coor, aes(x = long, y = lat, color = mean_res_time)) +
#  viridis::scale_color_viridis()

#nyc_pic = readJPEG("nyc-boroughs-map.jpg")

#map_plot = 
#  incident_zip_coor %>% 
#  ggplot(aes(x = long, y = lat, size = mean_res_time, color = mean_res_time)) +
#  annotation_custom(rasterGrob(nyc_pic, 
#                               width = unit(1,"npc"), 
#                               height = unit(1,"npc")), 
#                               -Inf, Inf, -Inf, Inf) +
#  geom_point() +
#  viridis::scale_color_viridis() + theme_set(theme_bw() + theme(legend.position = "bottom"))


#ggplotly(map_plot) 


prcp_eda =
  finaldat %>% 
  mutate(prcp_ctg = fct_relevel(prcp_ctg, c("no_prcp", "low", "high"))) %>%   group_by(date, prcp_ctg) %>% 
  summarise(mean_resp_time = mean(response_time)) %>% 
  ggplot(aes(x = prcp_ctg, y = mean_resp_time)) +
  geom_violin(aes(fill = prcp_ctg), alpha = .3) +
  stat_summary(fun.y = mean, geom = "point", size = 2, color = "blue") +
  labs(
    title = "Rainy conditions",
    y = "Mean response time(min)",
    x = " "
  ) +
  scale_x_discrete(labels = c("0(mm)", "0-25(mm)", "25(mm)+")) +
  viridis::scale_fill_viridis(
    name = "Precipitation",
    discrete = TRUE) +
  theme(plot.title = element_text(size = 12),
        axis.title.y = element_text(size = 8),
        axis.text.x = element_text(size = 9),
        legend.position = "None") 

snow_eda =
  finaldat %>% 
  mutate(snow_ctg = fct_relevel(snow_ctg, c("no_snow", "low", "high"))) %>%   group_by(date, snow_ctg) %>% 
  summarise(mean_resp_time = mean(response_time)) %>% 
  ggplot(aes(x = snow_ctg, y = mean_resp_time)) +
  geom_violin(aes(fill = snow_ctg), alpha = .3) +
  stat_summary(fun.y = mean, geom = "point", size = 2, color = "blue") +
  labs(
    title = "Snowy conditions",
    y = "Mean response time(min)",
    x = " "
  ) +
  scale_x_discrete(labels = c("0(mm)", "0-50(mm)", "50(mm)+")) +
  viridis::scale_fill_viridis(
    name = "Snow",
    discrete = TRUE) +
  theme(plot.title = element_text(size = 12),
        axis.title.y = element_text(size = 9),
        axis.text.x = element_text(size = 8),
        legend.position = "None") 

prcp_snow = prcp_eda + snow_eda 

ggplotly(prcp_snow)
--!>

